---
title: "Causal Group Project-Preliminary Project Description"
author: "Group Members: Shelley Facente, Steph Holm, Lizzy Kinnard, Veronica Pear"
date: "3/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Target Population/Inclusion Criteria of Study 
These data are from the National Longitudinal Survey of Youth 1997. This was recruited as a nationally representative cohort of youth age 12-16 (initial n=9000) in 1997. These youth have since been followed longitudinally. The target population is youth in the United States.

# Causal Question:  
=======
Does the experience of having been bullied prior to age 12 effect incidence of drug use in adolescence or adulthood?

# Target Causal Parameter: Lizzy

# Data Description: Veronica

```{r}
#sample size calculation

#marginal distribution of exposure

#marginal distribution of outcome
```


# Anticipated Challenges: Shelley
=======
There are many potential Ws in our dataset, and in order to make this manageable we will need to determine the appropriate variables to include in our analysis. We have already verified that there is sufficient data in each of the strata (as found in the data description) so while we were originally concerned about this, we think it will be fine. Our major challenge will likely be that this is longitudinal data, which we have only dealt with in small ways in class; we will need to pay careful attention to time-ordered variables and make sure they are appropriately handled in our causal model. 