---
title: "Causal Group Project-Preliminary Project Description"
author: "Group Members: Shelley Facente, Steph Holm, Lizzy Kinnard, Veronica Pear"
date: "3/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Target Population/Inclusion Criteria of Study (See nlsy97): Steph


# Causal Question:  
=======
Does the experience of having been bullied prior to age 12 effect incidence of drug use in adolescence or adulthood?

# Target Causal Parameter: Lizzy

# Data Description: Veronica

```{r}
#sample size calculation

#marginal distribution of exposure

#marginal distribution of outcome
```


# Anticipated Challenges: Shelley
=======
There are many potential Ws in our dataset, and in order to make this manageable we will need to determine the appropriate variables to include in our analysis. We have already verified that there is sufficient data in each of the strata (as found in the data description) so while we were originally concerned about this, we think it will be fine.